Looking for CDI Bias: WS Production

### 2021-08-25 03-preliminary-WSprod.Rmd
added WS production fits

---
title: "Looking for CDI Bias: WG Comprehension"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

@@

First, let's get some wordbank data for American English. We'll start with production data from the CDI: Words & Gestures (WG) form.
We'll look for DIF based on 1) sex, 2) SES (high vs. low), and 3) race/ethnicity. 
Questions: How do we choose a dozen anchor items (that we expect to be unbiased)?

```{r get-wordbank-data, eval=F, echo=F}
# Run Once
#get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

# 8-18 month-olds
table(d_demo$age)

#sum(is.na(d_demo$mom_ed)) # many mom_ed missing

```

Let's do comprehension data.

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(understands = as.numeric(value == "yes" | value == "produces" | value == "understands")) # we'll start with comprehension
# yes = understands; no = does not understand/produce, produces = produces (and understands)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word") %>% select(-complexity_category, -item_id)

# subjects not producing any words yet (can't fit model with them)
bad_Ss = d_demo[which(d_demo$comprehension==0),]$data_id # 8 bad comprehension Ss

d_long_wg <- d_long_wg %>% left_join(en_wg %>% select(num_item_id, definition)) %>%
  filter(!is.na(definition))

d_mat <- d_long_wg %>% select(data_id, definition, understands) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = understands)

d_long_wg <- d_long_wg %>% select(-value, -num_item_id)

save(d_mat, d_demo, en_wg, d_long_wg, file=here("data/en_wg_comp.Rdata"))
```

```{r load-data}
# load saved CDI:WG data
load(here("data/en_wg_comp.Rdata")) # d_demo, d_mat, en_wg (items)

d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id)

@@

@@

table(d_demo$mom_ed)
# high SES = mom_ed >= Some College (739 high, 324 low-SES)
d_demo <- d_demo %>% mutate(ses_group = ifelse(as.numeric(mom_ed)>4, "high", "low"))

table(en_wg$category) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We will fit comprehension data from `r nrow(d_demo)` children for the WG form.

@@

save(file = here("data/eng_wg_comp_mod_2pl.Rds"), "m1", "coefs_2pl", "fscores_2pl") # ,
```

@@

```{r fit-rasch-group-models, echo=F, eval=F}
# first fit Rasch model to each group, and then compare resulting parameters

m1pl_m <- mirt(d_mat[which(d_demo$sex=="Male"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)
m1pl_f <- mirt(d_mat[which(d_demo$sex=="Female"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)

m1pl_h <- mirt(d_mat[which(d_demo$ses_group=="high"),], 1, "Rasch", SE.type='Oakes', verbose=T)
m1pl_l <- mirt(d_mat[which(d_demo$ses_group=="low"),], 1, "Rasch", SE.type='Oakes', verbose=T)

save(file = here("data/eng_wg_rasch_group_models.Rds"),
     "m1pl_m", "m1pl_f", "m1pl_h", "m1pl_l")

res_m = residuals(m1pl_m)
#m1pl_m_ci = boot.mirt(m1pl_m)
#itemplot(m1pl_m, 1, CE=T)
```

```{r, echo=F, include=F}
load(here("data/eng_wg_rasch_group_models.Rds"))

# coef(m1pl_m) shows CI_2.5 and CI_97.5 for d...iterate over and add to dataframes?
#ci_lo = coef(m1pl_m)$`baa baa`['CI_2.5','d']
#ci_hi = coef(m1pl_m)$`baa baa`['CI_97.5','d']

par_m = as_tibble(coef(m1pl_m, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_m, simplify = TRUE)$items))
par_f = as_tibble(coef(m1pl_f, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_f, simplify = TRUE)$items))

par_h = as_tibble(coef(m1pl_h, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_h, simplify = TRUE)$items))
par_l = as_tibble(coef(m1pl_l, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_l, simplify = TRUE)$items))

rasch_pars = tibble(definition = par_m$definition,
                    d_Male = par_m$d,
                    d_Female = par_f$d,
                    d_LowSES = par_l$d,
                    d_HighSES = par_h$d) %>%
  mutate(d_sex_diff = abs(d_Female - d_Male),
         d_ses_diff = abs(d_HighSES - d_LowSES))

sex_thresh = median(rasch_pars$d_sex_diff) + sd(rasch_pars$d_sex_diff)
ses_thresh = median(rasch_pars$d_ses_diff) + sd(rasch_pars$d_ses_diff)

p1 <- ggplot(rasch_pars, 
       aes(x = d_Male, y = d_Female)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_sex_diff > sex_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')

p2 <- ggplot(rasch_pars, 
       aes(x = d_LowSES, y = d_HighSES)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_ses_diff > ses_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Low SES") + 
  ylab("Item Ease for High SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

```{r, echo=F, fig.width=10, fig.height=5, include=F}
ggarrange(p1, p2)
```

@@

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg3_dif, mg1, mg2, mg3, mg4, m_sc2, mod_sc1, mod_fullconstrain) 
```

```{r}
load(here("data/eng_wg_comp_sex_sexDIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

@@

mg2_dif = DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.
mg3_dif = DIF(mg3, c('d')) # (though varying group mean not justified according to AIC/BIC)

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg1, mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain,
     mg2_dif, mg3_dif)
```

>>>>

---
title: "Looking for CDI Bias: WS Production"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

@@

First, let's get some wordbank data for American English. We'll now turn to production data from the CDI: Words & Sentences (WS) form.
We'll look for DIF based on 1) sex, 2) SES (high vs. low), and 3) race/ethnicity.

```{r load-data}
# see GLIMMER_production.R
load(here("data/en_ws_production.Rdata"))

d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))

@@

# only 4090 kids with demographic variables

@@

sex_group = as.character(d_demo$sex)
ses_group = as.character(d_demo$ses_group)
eth_group = as.character(d_demo$eth_group)

table(d_demo$mom_ed)

table(en_ws$category) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We will fit comprehension data from `r nrow(d_demo)` children for the WS form.

@@

save(file = here("data/eng_ws_prod_mod_2pl.Rds"), "m1", "coefs_2pl", "fscores_2pl") 
```

@@

@@

save(file = here("data/eng_ws_prod_sex_sexDIF.Rds"), 
     mg3_dif, mg1, mg2, mg3, mg4, m_sc2, mod_sc1, mod_fullconstrain) 
```

```{r}
load(here("data/eng_ws_prod_sex_sexDIF.Rds"))
load(here("data/eng_ws_prod_mod_2pl.Rds"))

@@

#residuals(mod_sc1)

mg2_dif = DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.
#mg3_dif = DIF(mg3, c('d')) # (though varying group mean not justified according to AIC/BIC)

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg1, mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain,
     mg2_dif) # mg3_dif
```
